Is the direction of Granger causal influence the same as the direction of information flow?

### 2015-07-07 paper.tex
Make corrections to the estimating coeffs section.

The Granger Causality and Directed Information metrics are a function of the regression coefficients estimated from the data by fitting the models given by equations~\ref{eq:ar-model-x-x} and~\ref{eq:ar-model-x-x-theta}. In our analysis, we described the data-generation model: the algorithm inspired by the Schalkwijk and Kailath scheme for feedback communication. We then proceeded to use the regression coefficients generated by this model directly. This could be construed as erroneous: we ought to simulate the data generation, and compute the regression coefficients from the generated data. This would better model the actions of the neuroscientist who seeks to perform a Granger Causal analysis.

We justify our use of the theoretical regression coefficients as follows: we assume that the regression coefficients can be accurately estimated from data, since neuroscientific experiments typically record the \emph{same} processes multiple times -- these are called ``trials''. The availability of multiple trials of the same process can be leveraged to estimate the regression coefficients accurately, even if the processes are non-stationary.

@@

If we call the vector on the LHS $\bf{Y}$ and the matrix on the RHS $\bf{X}$, then the vector of $\alpha_j(t)$'s ($\bf{\alpha}(t)$) can be estimated at time instant $t$ using Ordinary Least Squares:
\begin{equation*}
\hat{\bf{\alpha}}(t) = (\bf{X}^T \bf{X})^{-1} \bf{X}^T \bf{Y}
\end{equation*}
This is an unbiased and consistent estimator for $\bf{\alpha}(t)$. This analysis can be trivially extended to the model described by equation~\ref{eq:ar-model-x-x-theta}. With a sufficiently large number of trials, therefore, the regression coefficients can be estimated to arbitrarily high accuracy.

As a final point, we note that estimating the regression coefficients accurately is a form of worst-case analysis. As mentioned at the end of section~\ref{sec:previous-criticisms-of-gc}, we see that \emph{despite} being computed accurately, the metrics of Granger Causality and Directed Information incorrectly estimate the direction of information flow.

>>>>

The Granger Causality and Directed Information metrics are a function of the regression coefficients estimated from the data by fitting the models given by equations~\ref{eq:ar-model-x-x} and~\ref{eq:ar-model-x-x-theta}. In our analysis, we described the data-generation model: the algorithm inspired by the Schalkwijk and Kailath scheme for feedback communication. We then proceeded to use the system parameters of this model directly as regression coefficients in our Granger Causality computation. This could be construed as being erroneous: we ought to simulate the data generation, and estimate the regression coefficients from the generated data. This would better model the actions of the neuroscientist who seeks to perform Granger Causality analysis.

We justify our knowledge of the system parameters and their use as regression coefficients in the following manner: we assume that the regression coefficients can be accurately estimated from data, since neuroscientific experiments typically record the \emph{same} processes multiple times -- these are called ``trials''. The availability of multiple trials of the same process can be leveraged to estimate the system parameters accurately, even if the processes are non-stationary.

@@

If we call the vector on the LHS $\bf{Y}$ and the matrix on the RHS $\bf{X}$, then the vector of $\alpha_j(t)$'s ($\bf{\alpha}(t)$) can be estimated at time instant $t$ using Ordinary Least Squares:
\begin{equation*}
\hat{\bf{\alpha}}(t) = (\bf{X}^T \bf{X})^{-1} \bf{X}^T \bf{Y}
\end{equation*}
This is an unbiased and consistent estimator for $\bf{\alpha}(t)$. This analysis can be trivially extended to the model described by equation~\ref{eq:ar-model-x-x-theta}. With a sufficiently large number of trials, therefore, the system parameters (to be used as regression coefficients in the Granger Causality analysis) can be estimated to arbitrarily high accuracy.

As a final remark, we note that estimating the regression coefficients accurately is a conservative assumption on our part. As mentioned at the end of section~\ref{sec:previous-criticisms-of-gc}, we see that \emph{despite} being computed accurately, the metrics of Granger Causality and Directed Information incorrectly estimate the direction of information flow. A rigorous analysis would warrant the computation of these coefficients in simulation, for a finite number of trials. It is our belief, however, that our result is unlikely to degrade if the coefficients are not estimated perfectly. Future work will address this matter in greater depth.
